CALVIS: chest, waist and pelvis circumference from 3D human body meshes as ground truth for deep learning

### 2019-10-30 PaperRMFI2019/Paper_ID_25_Draft.tex
anonym submission

\begin{document}
%
\title{CALVIS: chest, waist and pelvis circumference from 3D human body meshes	
as ground truth for deep learning}
%
\titlerunning{Chest, wAist and peLVIS circumference from 3D human body meshes}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Yansel Gonzalez Tejeda\orcidID{0000-0003-1002-3815} \and
Helmut Mayer\orcidID{1111-2222-3333-4444}}
%
\authorrunning{Gonzalez-Tejeda and Mayer}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Paris Lodron University of 
	Salzburg, Kapitelgasse 4-65020 Salzburg, Austria
\url{https://www.uni-salzburg.at}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper we present CALVIS, a method to calculate \textbf{C}hest, 
w\textbf{A}ist and pe\textbf{LVIS} circumference from 3D human body meshes. Our 
motivation is to use 
this data as ground truth for training convolutional neural networks 
(CNN). Previous work had used the large scale CAESAR dataset or determined 
these anthropometrical measurements $\textit{manually}$ from a person 
or human 3D body meshes. 
Unfortunately, acquiring these data is a cost and 
time consuming endeavor. In contrast, our method can be used on 
3D meshes automatically. We synthesize eight human body meshes and 
apply 
CALVIS to calculate chest, waist and pelvis circumference. We evaluate the 
results qualitatively and observe that the measurements can indeed be used 
to estimate the shape of a person. We then asses the 
plausibility of our approach by generating
ground truth with CALVIS to train a small CNN. After
having trained the network with 
our data, we achieve competitive validation error.
Furthermore, we make the implementation of CALVIS publicly available to 
advance the field.

@@

We conduct two experiments. In the first experiment we synthesize eight (four 
female and four male) human body meshes using shape parameters provided by 
SURREAL \cite{varol17_surreal}.
The meshes reflect human figures characteristics such as bulky, slim, small and 
tall subjects. Then we apply CALVIS to calculate chest, waist and pelvis 
circumference. Since we do not have ground truth, we evaluate the results 
qualitatively.
The second 
experiment serves to asses the plausibility of our approach to use the 
synthetic data for deep learning. We input the calculated human 
dimensions to an artificial neural network. After having trained the network 
with our data, we show that the network is able to learn the task.

\subsection{Qualitative evaluation}\label{subsec:qualitative_eval}
Here we evaluate our method qualitatively.
\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{experiment_1_results.png}
	\end{center}
	\caption{Qualitative evaluation for male and female meshes.}
	\label{fig:qualitative_eval}
\end{figure}

Figure \ref{fig:qualitative_eval} shows the calculated dimensions on eight 
subjects, male subjects no. 1-4 and female subjects no. 5-8. The black paths 
represent chest, waist and pelvis circumference (from 
top to bottom) as defined by equations \ref{eq:cc}-\ref{eq:pc}, respectively. 
We slightly rotate the meshes and add transparency for visualization purposes. 
Note that our method is able to estimate automatically the human body 
dimensions for all 
subjects. One particularity can be observed regarding the chest circumference: 
while for all male and female subjects except for subject no. 7 the chest 
circumference is 
estimated above the breast, for subject no.7 the chest circumference is 
estimated around the breast.

\subsection{Learning with Calvis-Net}\label{subsec:learning}
This experiment demonstrates the plausibility of employing CALVIS to obtain 
data 
that can be used as a ground truth for machine learning (in general) and 
specifically as supervising signal for CNN.

\subsubsection{Experimental setup} We annotate the CMU dataset with CALVIS. The 
automatic annotation process took approximately $1\,h\, 47\, min$ in an 
enhanced 
modern personal computer.

\subsubsection{Calvis-CMU Dataset} We synthesize x images of every 
person using the package Blender. To increase the realism of these images we 
adopt a yansel lighting model. This makes sense because it resembles the way 
how a person is standing in a room when she/he goes to the tailor for her/his 
body dimensions to be taken.

\subsubsection{Synthetic images as input} We synthesize x images of every 
person using the package Blender. To increase the realism of these images we 
adopt a yansel lighting model. This makes sense because it resembles the way 
how a person is standing in a room when she/he goes to the tailor for her/his 
body dimensions to be taken.

@@

Figure \ref{fig:calvis_net} shows our prototypical CNN Calvis-Net where the 
input is a synthetic image from a 3D mesh and the output are three body 
dimensions.

Figure \ref{fig:learning_curve} shows that after epoch number 5 the estimation 
error decreases and the learning converges.

We achieve mean absolute validation error of 0.061698 m plus +-.

\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{Training_val_loss.png}
	\end{center}
	\caption{Learning convergence evaluation.}
	\label{fig:learning_curve}
\end{figure}

%------------------------------------------------------------------------
\section{Conclusion}
In this paper we presented CALVIS, a method to calculate chest, waist and pelvis circumference from 3D human meshes. 
We demonstrate that our approach can be used to train a CNN by imputing synthetic images of humans and using the measurements calculated with CALVIS as supervision signal.
Furthermore, we contribute with a prototype CNN CALVIS-NET. Our experiments show that the CNN is able to learn how to perform this task, achieving mean absolute validation error of 0.061698 m plus +-.
The code 
and data use 
in this paper will be made publicly available for researchers.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{egbib}
%

>>>>

\begin{document}
%
\title{CALVIS: chest, waist and pelvis circumference from 3D human body meshes	
as ground truth for deep learning}
%
\titlerunning{Chest, waist and pelvis circumference from 3D human body meshes}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Author1\orcidID{0000-0000-0000-0000} \and
Author2\orcidID{0000-0000-0000-0000}}
%
\authorrunning{Author1 and Author2}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Institute1}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper we present CALVIS, a method to calculate \textbf{C}hest, 
w\textbf{A}ist and pe\textbf{LVIS} circumference from 3D human body meshes. Our 
motivation is to use 
this data as ground truth for training convolutional neural networks 
(CNN). Previous work had used the large scale CAESAR dataset or determined 
these anthropometrical measurements $\textit{manually}$ from a person 
or human 3D body meshes. 
Unfortunately, acquiring these data is a cost and 
time consuming endeavor. In contrast, our method can be used on 
3D meshes automatically. We synthesize eight human body meshes and 
apply 
CALVIS to calculate chest, waist and pelvis circumference. We evaluate the 
results qualitatively and observe that the measurements can indeed be used 
to estimate the shape of a person. We then asses the 
plausibility of our approach by generating
ground truth with CALVIS to train a small CNN. After
having trained the network with 
our data, we achieve competitive validation error.
Furthermore, we make the implementation of CALVIS publicly available to 
advance the field.

@@

We conduct two experiments. In the first experiment we synthesize eight (four 
female and four male) human body meshes using shape parameters provided by 
SURREAL \cite{varol17_surreal}.
The meshes reflect human figures characteristics such as bulky, slim, small and 
tall subjects. Then we apply CALVIS to calculate chest, waist and pelvis 
circumference. Since we do not have ground truth, we evaluate the results 
qualitatively.
The second 
experiment serves to asses the plausibility of our approach to use the 
synthetic data for deep learning.

\subsection{Qualitative evaluation}\label{subsec:qualitative_eval}
A qualitative evaluation is in this case pertinent. There is currently no benchmark to measure how accurate the chest, waits and pelvis circumference can be. The main issue is to identify \textit{where} these dimensions are located in the body. For instance, even domain experts like tailors and anthropometrists can provide human body dimensions with limited precision.
\begin{figure}[H]
	\begin{center}
		%\fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
		\includegraphics[width=\linewidth]{experiment_1_results.png}
	\end{center}
	\caption{Qualitative evaluation for male and female meshes.}
	\label{fig:qualitative_eval}
\end{figure}

Figure \ref{fig:qualitative_eval} shows the calculated dimensions on eight 
subjects, male subjects no. 1-4 and female subjects no. 5-8. The black paths 
represent chest, waist and pelvis circumference (from 
top to bottom) as defined by equations \ref{eq:cc}-\ref{eq:pc}, respectively. 
We slightly rotate the meshes and add transparency for visualization purposes. 
Note that our method is able to estimate automatically the human body 
dimensions for all 
subjects. The dimensions are indeed estimated at locations where domain experts regularly identify them.

\subsection{Learning with Calvis-Net}\label{subsec:learning}
This experiment demonstrates the plausibility of employing CALVIS to obtain 
data 
that can be used as a ground truth for machine learning (in general) and particularly as supervising signal for CNN.

\subsubsection{Experimental setup.} First, we assemble our Calvis-CMU Dataset. We synthesize 3400 grayscale images (1700 of each gender) of human body meshes generated with parameters from the CMU dataset using Blender. To increase the realism of these images we 
adopt an indoor lighting model with one light source and white background. This makes sense because it resembles the way 
how a person is standing in a room when she attends to the tailor for her 
body dimensions to be taken.
Subsequently, we annotate the CMU dataset with CALVIS. The 
automatic annotation process took approximately $1\,h\, 47\, min$ in an 
enhanced modern personal computer.

@@

Figure \ref{fig:calvis_net} shows our prototypical CNN Calvis-Net where the 
input is a synthetic image from a 3D human mesh and the output are three body 
dimensions. Implemented with pytorch \cite{paszke2017automatic}. We train the network for 20 epochs and perform 3-fold cross-validation  with skorch. We achieve mean absolute validation error of 12 mm.

%------------------------------------------------------------------------
\section{Conclusion}
In this paper we presented CALVIS, a method to calculate chest, waist and pelvis circumference from 3D human meshes. 
We demonstrate that our approach can be used to train a CNN by imputing synthetic images of humans and using the measurements calculated with CALVIS as supervision signal.
Furthermore, we contribute with a prototype CNN CALVIS-NET. Our experiments show that the CNN is able to learn how to perform this task, achieving competitive mean absolute validation error.
The code 
and data use 
in this paper will be made publicly available for researchers.
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{egbib}
%
